A Gambler's DAgger: Reducing Supervisor Burden in Online Learning From Demonstration Learning Through Hierarchical Supervisors

### 2016-02-22 paper.tex
change stuff

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\usepackage[ruled,vlined]{algorithm2e}

@@

\renewcommand{\baselinestretch}{.92}

%\title{Iterative Imitation Learning with Reduced Human Supervision [v11]}
%\title{SHIV:  Reducing Human Supervision for Robot Active Learning [v11]}

\title{SHIV: Reducing Supervisor Burden using Support Vectors for Efficient
Learning from Demonstrations in High Dimensional State Spaces }

\author{Michael Laskey$^1$, Sam Staszak$^1$, Wesley Yu-Shu Hsieh$^1$, Jeffrey Mahler$^1$, \\ Florian T. Pokorny$^1$, Anca D. Dragan$^1$, and Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh,ftpokorny,anca\}@berkeley.edu, \small staszass@rose-hulman.edu} }%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

\begin{abstract}
Online learning from demonstration algorithms, such as DAgger, can learn policies for problems where the system
dynamics and the cost function are unknown. However, during learning, they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states while executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936 dimension visual HOG feature state representation, 2) a push-grasping in clutter simulation for a 4DOF robot arm, using Box2D, with a 22 dimension state representation  of joint angles and the pose of each object, and 3) a surgical needle insertion task with a 16 dimension state represented spatial transform.  Results suggest that SHIV can learn equivalent policies with up to 70\% fewer queries.

 \end{abstract}

@@

\section{Introduction}

In model-free robot Learning from Demonstration (LfD), a robot learns to perform a task, such as driving or grasping an object in a cluttered environment, from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to either the cost function that it should optimize, nor the dynamics model. The former occurs when it is difficult to specify how to trade-off various aspects that matter, like a car trying to drive on the road and dodge other cars \cite{abbeel2004apprenticeship}.  The latter occurs when either the system or the interaction with the world is difficult to characterize like when a robot is trying to grasp an object in clutter and doesn't have an accurate model of the objects dynamics.

Rather than explicitly learning the cost function, like in Inverse Reinforcement Learning \cite{ng2000algorithms}, and the dynamics model, and then using optimization to produce a policy for the task, in model-free LfD the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}. Learning from demonstration has been used successfully in recent year for a large number of robotic tasks, including helicopter maneuvering \cite{abbeel2007application},car parking \cite{abbeel2008apprenticeship},  and robotic surgery \cite{van2010superhuman}.

\begin{figure}[ht]
\centering

\includegraphics[width=0.5\textwidth]{figures/teaser.pdf}
\caption{ \footnotesize
SHIV  reduces the number of queries to the supervisor in our three test domains: a) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; b) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; c) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery; d) normalized performance of SHIV and DAgger's policy on the grasping in clutter example versus states labeled by the supervisor.
}
\vspace*{-20pt}
\label{fig:teaser}
\end{figure}

LfD algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.
In offline LfD, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will eventually put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

Online LfD addresses this issue by iteratively gathering more examples from the supervisor in states the robot encounters~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. One such algorithm, DAgger, learns a series of policies. At each iteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy, and the supervisor provides demonstrations for all states the robot visits. The new state/control examples are aggregated with the old examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, under certain conditions, the number of errors scales only linearly with the time horizon of the task\cite{ross2010reduction}.

However one drawback is that DAgger significantly increases the burden on the supervisor, who now labels all states that the robot visits during training.  Our goal is to reduce supervisor burden without degrading performance.

Our key insight is that the robot only needs to request demonstrations for risky states. Risk arises because:  (1) states are sufficiently different than the states previously encountered, i.e., they can be considered novel or outliers~\cite{hodge2004survey}; or (2) states are in a region that has been misclassified in previous training data.
%\adnote{Ken's version below does not seem to represent what we actually do.}
%the action associated with a state under the current policy yield a resulting state that is inconsistent with the expected next state. [how do we measure 2? - Ken]

We present an algorithm for online LfD that can actively decide whether it is necessary to ask the supervisor to label the new states that the robot encounters. SHIV  (Svm-based reduction in Human InterVention) reduces supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to the non-stationary state distributions the robot encounters as it iteratively executes learned policies. We build on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

\adnote{nothing about experiements and resutls? how do i know if it worked? also how do i know that you didn;t just compare to dagger, but ran an indepth analysis of the algorithm? i thought i put a paragraph like that here at some point}

\section{Related Work}
Below we summarize related work in active approaches to robot LfD and risk evaluation techniques.

\noindent\textbf{Active Learning from Demonstration}
We formulate applying active learning to Online Lfd as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a label or control signal) or not is not over the entire state space (like in traditional pool-based active learning), but on states drawn one at a time from some data stream. In  online LfD, this data stream is the states encountered when the robot is executing the current best policy.

Aside from the data stream, stream-based active learning has another ingredient: a query selection method. A query selection method is deciding whether or not to query. Typical query selection methods are estimator-centric, evaluating risk using the estimator, e.g. distance from the classification hyperplane \cite{tong2002support}, as in Fig. \ref{fig:support_example}(a); or query by committee  \cite{breiman1996bagging}, which uses a committee of hypothesized estimators that are trained on different subsets of the training data. Risk in query by committee is based on the level of agreement or disagreement among these hypotheses, with higher levels of disagreement for a state leading to higher chances of querying that state (Fig. \ref{fig:support_example}(b)).

Both approaches implicitly assume a stationary state distribution --- that the new data is sampled from the same distribution as the previous training data. Although such methods have been previously proposed for online LfD (see \cite{chernova2009interactive,grollman2007dogged} for the former and \cite{judah2011active,judah2012active} for the latter),
online LfD violates the stationary distribution assumption because each new policy induces a new distribution of states. This can have negative consequences for learning: it has been shown that when training and test distributions are different, query by committee can perform worse than randomly selecting  which states to query \cite{burbidge2007active}.

The problem lies in being estimator-centric. The  estimator is a function of the current distribution. Therefore, it no longer provides a good measure of confidence when that distribution changes, as it does when the robot is rolling out a learned policy and starts encountering further away states. Instead of relying on the estimator, our measure of risk explicitly identifies when states are drawn from a different distribution, i.e. when they are novel.

%%In SHIV, we use a query selection method that allows for a non-stationary distribution, evaluating risk based on novelty (is the state an outlier to the training distribution), as in Fig. \ref{fig:support_example}(c). Additionally, we also assign higher risk to  historically mislabeled regions, as in Fig. \ref{fig:support_example}(d). We then use a classification method based on the One Class SVM \cite{scholkopf2001estimating} to predict if a state is risky or not risky, which we we define as safe.%Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}.

\noindent\textbf{Risk via Novelty Detection}
The main motivation for our risk definition is the notion that a trained model will be able to generalize within the distribution it is
trained on \cite{tokdar2010importance}, making states outside of this distribution risky. Novelty detection \cite{hodge2004survey} is concerned with recognizing when this happens: recognizing that a sample is outside of this distribution.

Our work builds on Kim et al. \cite{kim2013maximum} who apply novelty detection for safe online learning from demonstrations to enable a supervisor to take over and prevent unsafe states.  In their MMD-LI algorithm, risk is evaluated with Maximal Mean Discrepancy, which for a single point is the sum of a kernel density estimate and variance on the dataset, which works well in low dimensional state spaces.  MMD-LI also reduces supervisor burden because the supervisor only takes control and provides demonstrations in risky states. However, having the supervisor take over can bias the states sampled, which can in practice lead to a loss in performance \cite{ross2010reduction}.  In contrast, we focus on active learning for online LfD for high-dimensional state spaces, do not assume the supervisor is always available, and propose an alternative to non-parametric density estimation which can require data exponential in the dimension of the state-space
\cite{nadaraya1964estimating}.

An alternative to evaluating the probability of a data point is to measure distance to its nearest neighbors \cite{knox1998algorithms}. However, this approach was shown to be
susceptible to issues, since nearest neighbors  incorporates only local information about the data. For example, a group of outliers can be close together, but significantly far from the majority of the data and nearest neighbors would mark them as not novel \cite{hodge2004survey}.

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to the find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge, incorporating a query selection method for nonstationary distributions, such as the One Class SVM, into online LfD is a novel contribution.

\section{Problem Statement}
The goal of this work is to learn a policy that matches that of the supervisor's while asking the supervisor for as few examples as possible.

@@

We model the initial state as sampled from a distribution over the state space.
We assume a known state space and set of controls. We also assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a supervisor who can, given a state, provide a control signal label. We additionally assume the supervisor can be noisy and imperfect, noting that the robot cannot surpass the level of performance of the supervisor.

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of 
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

A policy is a function $\pi: \mathcal{X} \to \mathcal{U}$ from states to control inputs. 
We consider a space of policies $\pi_{\theta}:\mathcal{X}\to \mathcal{U}$ parameterized by some $\theta\in \mathbb{R}^d$. Any such policy $\pi_{\theta}$ in an environment with probabilistic initial state density and Markovian dynamics
induces a density on trajectories of length $T+1$: $$p(\tau | \theta)=
p(\bx_0)\prod_{i=0}^{T-1}p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$$

Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ from time $0$ to time $t-1$.  Following \cite{ross2010reduction}, we can compute
the average density on states for any timepoint by $p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)$.

@@

\noindent\textbf{Objective.} The objective of  policy learning is to find a policy that minimizes some known cost function $C(\hat{\tau}) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\hat{\tau}$. The cost $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in the task of inserting a peg into a hole, a function on distance between the peg's current and desired final state is used \cite{levine2015end}.

In our problem, we do not have access to the cost function itself. Instead, we only have access to 
a supervisor that can achieve a desired level of performance on the task. The supervisor provides the robot an initial set
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training data set $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

We define a `surrogate' loss function as in \cite{ross2010reduction}, $l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2_2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

@@

Following \cite{ross2010reduction}, our objective is to find a policy $\pi_{\theta}$ minimizing the expected surrogate loss, where the expectation is taken over the distribution of states induced by the policy across any time point in the horizon:

@@

 If the robot could learn the policy  perfectly, this state density would match the one encountered in user examples. But if the robot makes an error, that error changes the distribution of states that the robot will visit, which can lead to states that are far away from any examples and difficult to generalize to ~\cite{pomerleau1989alvinn}.This motivates iterative algorithms like DAgger, which iterate between learning a policy and the supervisor providing feedback. The feedback is in the form of control signals on states sampled from the robot's new distribution of states.

\section{Risky and Safe States}

@@

\includegraphics[width=\textwidth]{figures/active_learning.pdf}

>>>>

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{url}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

@@

\renewcommand{\baselinestretch}{.95}
\usepackage{times}
\usepackage{microtype}
%\title{Iterative Imitation Learning with Reduced Human Supervision [v11]}
%\title{SHIV:  Reducing Human Supervision for Robot Active Learning [v11]}

\title{A Gambler's DAgger: Reducing Supervisor Burden in Online Learning From Demonstration Learning  Through Hierarchical Supervisors}

\author{Michael Laskey$^1$, Jonathan Lee$^1$, Caleb Chuck$^1$, David Gealy$^1$,...,Florian T. Pokorny,\\
 Anca D. Dragan$^1$, and Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh,ftpokorny,anca\}@berkeley.edu, \small staszass@rose-hulman.edu} }%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

\begin{abstract}
Online learning from demonstration algorithms such as
DAgger can learn policies for problems where the system dynamics and
the cost function are unknown. However they impose a
burden on supervisors to respond to queries each time the robot
encounters new states while executing its current best policy.  Traditionally there has only been one supervisor and it is deemed expert. However, an expert supervisor might not be needed for a large portion of the state space, which could be handled by less-skilled supervisors. We present the Gambler's DAgger, which instead of only using an expert supervisors presents a hierarchical supervisor model, which first leverages cheaper supervisors, such as analytical methods on approximate dynamics or crowdsourced Amazon Mechanical Turk workers to train the policy and then only consults an expert for harder to learn states. We demonstrate this approach in training a visuomotor deep network policy for 2D grasping in clutter on a 4-DOF Zymark Robot. We were able to reduce the number of queries to the final expert supervisor by $X$ and manage to collect $X$K examples by leveraging the hierarchical supervisor model. 
 \end{abstract}

@@

\section{Introduction} 
In model-free robot Learning from Demonstration (LfD), a robot learns to perform a task, such as driving or grasping an object in a cluttered environment, from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to either the cost function that it should optimize, nor the dynamics model. The former  when it is difficult to specify how to trade-off various aspects that matter, like a car trying to drive on the road with other cars~\cite{abbeel2004apprenticeship}.  The latter occurs when either the system or the interaction with the world is difficult to characterize, like when a robot is trying to grasp an object in clutter and does not have an accurate model of the object dynamics.

Rather than explicitly learning the cost function as in Inverse Reinforcement Learning and the dynamics model, and then using optimization to produce a policy for the task, in model-free LfD the robot learns a policy directly from supervisor examples, mapping states to controls~\cite{argall2009survey}. Learning from demonstration has been used successfully in recent year for a large number of robotic tasks, including helicopter maneuvering~\cite{abbeel2007application}, car parking~\cite{abbeel2008apprenticeship},  and robot surgery~\cite{van2010superhuman}.

 \begin{figure}
\includegraphics[width=0.5\textwidth]{f_figs/teaser.pdf}
\caption{Typical Initial States for our grasping in clutter task with a Zymark Robot. Red shapes indicate the clutter objects and the yellow circle indicates the goal object. The robot learns from raw image date to push the shapes and reach the yellow circle through our algorithm A Gambler's DAgger. The inscribed circle in the workspace prevents the robot from learning the trivial solution of simply pushing the red objects off the table.}
\vspace*{-20pt}
\label{fig:teaser}
\end{figure}

LfD algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.
In offline LfD, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small control error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will eventually put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

Online LfD addresses this issue by iteratively requesting more examples from the supervisor in states the robot encounters~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. One such algorithm, DAgger, learns a series of policies. At each iteration, the robot computes a policy based on the existing examples, then rolls out (executes) that policy, and the supervisor provides demonstrations for all states the robot visits. The new state/control examples are aggregated with the prior examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, under certain conditions, the number of errors scales only linearly with the time horizon of the task~\cite{ross2010reduction}.

One drawback is that DAgger significantly increases the burden on the supervisor, who must label all states that the robot visits during training.  Traditionally there has only been one supervisor and it is deemed an expert ~\cite{ross2010efficient,ross2010reduction,ross2013learning,duvallet2013imitation}. However, an expert supervisor might not be needed for a large portion of the state space, which could be handled by less-skilled supervisors.

Our key insight is that for a large portion of the task the supervisor's time is used teaching the robot something that a less costly supervisor could do. For instance in our grasping in clutter example, the robot spends a large portion of the time trying to learn how to go towards the goal object. Analytical techniques such as forward kinematics, template matching and motion planning are readily available to provide examples in these parts of the state space. However, traditional Online Lfd method exploit the expert supervisor for all parts of the state space.

We propose bootstrapping a policy from a hierarchy of supervisors ranked by cost. Examples of such a hierarchy are analytical models with a simplified dynamics of the world, crowd-source Amazon Mechanical Turk workers and then finally an expert supervisor. We demonstrate this approach in training a deep neural network for the visuo-motor based task of 2D grasping in clutter on a Zymark Robot. Our robot consists of an 3DOF arm that lies in a planar workspace and the ability to rotate the table. The objects dataset consists of 20 Medium Density Fiber extruded polygons taken from 2D projections of objects in the Dex-Net \cite{} database.  Results show that by bootstrapping from less skilled experts, the total amount of burden on the expert supervisor is reduced and the task is successfully learned \mlnote{We need an objective measure of success}.

\section{Related Work}
Below we summarize related work in reducing the number of queries to an expert in the Online LfD setting.\mlnote{Still a work in progress}

\noindent \textbf{Online Lfd with an Expert Supervisor}
Successful robotic examples of Online Learning From Demonstration where an expert have been employed have been in flying a quad-copter through a foster, navigating a wheel chair across a room and teaching a robot to follow verbal instructions and surgical needle insertion \cite{ross2013learning, kim2013maximum, duvallet2013imitation, laskey}.

However, to date these approaches have all used only an expert supervisor to label all parts of the state space. Our contribution consists of the use of a hierarchical structure of non-expert supervisors to reduce queries to the expert at the top of the hierarchy.

\noindent\textbf{Reducing Supervisor Burden in Online LfD} One approach that has been studied to reducing supervisor burden is applying active learning to only ask for queries one the robot is uncertain about the correct control to apply. Traditional active learning techniques like query-by-committee and uncertainty sampling have been applied by \cite{chernova2009interactive,judah2011active,grollman2007dogged}

Kim et al. demonstrated that due to the non-stationarity of the problem traditional active learning techniques do not work because the underlying state distribution changes and thus the use of novelty detection is needed to be more conservative in determining uncertainty~\cite{kim2013maximum}.

Laskey et al. created an approach, known as SHIV, for using active learning tailored to high dimensional and non-stationarity state distributions and  provided a modified version One Class SVM to reduce the problem to regularized binary classification~\cite{laskey}. However, the result that novelty detection is needed to reduce supervisor burden via active learning implies that for problems with high stochasticity, such as grasping in clutter with a large number of objects, the expert would be queried a significant number. Furthermore, with the advent of deep learning where the feature space changes each iteration, it becomes unclear as to what feature space to provide novelty detection in.

\noindent\textbf{Training Policies Via Analytical Methods} Instead of relying on a costly human supervisor an alternative approach is to use knowledge of the underlying dynamics and train a policy via a computationally expensive analytical method such as a controller in a physics simulator~\cite{IGOR} or dynamic programming in a Atari video game~\cite{NIPS2014_5421}.

While this approach can "compile" the computationally expensive option into an quickly computed function, it can be limited when analytical models cannot explicitly model the environment. For instance in our grasping in clutter example, the robot has no knowledge of the mass or friction properties of the object, which would make this a challenging problem for such analytical methods. We thus propose using analytical methods not as the expert supervisor, but instead part of a hierarchy where they can be used to reduce the burden of the supervisors above them in the hierarchy.

\section{Problem Statement}
The goal of this work is to learn a policy that matches that of an expert supervisor's while asking expert supervisor with fewer queries than DAgger.

@@

We model the initial state as sampled from a distribution over the state space.
We assume a known state space and set of controls. We also assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a set of supervisors who can, given a state, provide a control signal label. We additionally assume the supervisors can be noisy and imperfect, noting that the robot cannot surpass the level of performance of the current supervisor it is training with.

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of 
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$ that can be sampled from.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

A policy is a function $\pi: \mathcal{X} \to \mathcal{U}$ from states to control inputs. 
We consider a space of policies $\pi_{\theta}:\mathcal{X}\to \mathcal{U}$ parameterized by some $\theta\in \mathbb{R}^d$. Any such policy $\pi_{\theta}$ in an environment with probabilistic initial state density and Markovian dynamics
induces a density on trajectories. Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ from time $0$ to time $t-1$.  Following~\cite{ross2010reduction}, we can compute
the average density on states for any timepoint by $p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)$.

@@

\noindent\textbf{Objective.} The objective of  policy learning is to find a policy that maximizes some known cumulative reward function $\sum^T_{t=1} r(\bx_t,\bu_t)$ of a trajectory $\hat{\tau}$. The reward $r:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in the task of inserting a peg into a hole, a function on distance between the peg's current and desired final state is used~\cite{levine2015end}.

While a reward function exists for a task like grasping in clutter, such as a positive value for successfully reaching the goal object. It is often too delayed to provide useful information for quickly learning \cite{kitaevphysics}. We thus assume access to 
a supervisor that can achieve a desired level of performance on the task. The supervisor provides the robot an initial set
 of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying their policy. This induces a training data set $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

We define a `surrogate' loss function as in~\cite{ross2010reduction}, $l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2_2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

@@

Following~\cite{ross2010reduction}, our objective is to find a policy $\pi_{\theta}$ minimizing the expected surrogate loss, where the expectation is taken over the distribution of states induced by the policy across any time point in the horizon:

@@

 If the robot could learn the policy  perfectly, this state density would match the one encountered in user examples. But if the robot makes an error, that error changes the distribution of states that the robot will visit, which can lead to states that are far away from any examples and difficult to generalize to~\cite{pomerleau1989alvinn}. This motivates iterative algorithms like DAgger, which iterate between learning a policy and the supervisor providing feedback. The feedback is in the form of control signals on states sampled from the robot's new distribution of states.

 \section{A Gambler's DAgger}
 In this section, we first recall the original DAgger algorithm. Then introduce the notion and formal definition of a hierarchical supervisor and finally describe how to combine DAgger with a hierarchical supervisor to create a Gambler's DAgger.

 \subsection{DAgger: Dataset Aggregation}
 DAgger~\cite{ross2010reduction} solves the minimization in Eq. \ref{eq:LFD_obj} by iterating two steps: 1) compute a $\theta$ using the training data $\mathcal{D}$ thus far, and 2) execute the policy induced by the current $\theta$, and ask for labels for the encountered states.

\subsubsection{Step 1}
The first step of any iteration $k$ is to compute a $\theta_k$ that minimizes surrogate loss on the current dataset $\mathcal{D}_k=\{(x_i,u_i)|i\in\{1,\ldots,M\}\}$ of demonstrated state-control pairs (initially just the set $\mathcal{D}$ of initial trajectory demonstrations):

 \vspace{-1ex}
\begin{align}\label{eq:super_objj}
\theta_{k} = \underset{\theta}{\argmin} \: \sum_{i=1}^{M} l(\pi_{\theta}(\bx_i),\bu_i).
\end{align}

This sub-problem is a supervised learning problem, solvable by estimators like a support vector machine or a neural net. Performance can vary though with the selection of a the estimator~ Selecting the correct function class depends on the task being consider and knowledge of the problem, see for a guide~\cite{scholkopf2002learning}.

 \subsubsection{Step 2}
The second step  DAgger rolls out their policies, $\pi_{\theta_{k}}$, to sample states that are likely under $p(\bx|\theta_{k})$.  For every state visited, DAgger requests the supervisor to provide the appropriate control/label. Formally, for a given sampled trajectory  $\hat{\tau} = (\bx_0,\bu_0,...,\bx_T,\bu_T )$, the supervisor provides labels $\tilde{\bu}_t$, where $\tilde{\bu}_t \sim \tilde{\pi}(\bx_t) + \epsilon$, where $\epsilon$ is a small zero mean noise term, for $t\in \{0, \ldots, T\}$.
The states and labeled controls are then aggregated into the next data set of demonstrations $\mathcal{D}_{k+1}$:
$$D_{k+1}=\mathcal{D}_k \cup \{(\bx_t,\tilde{\bu_t})\|t\in\{0,\ldots,T\}\} $$
%\fpnote{terms such as aggregated etc would be much clearer if we just refer to a particular line in the algorithm}

Steps 1 and 2 are repeated for $K$ iterations or until 
the robot has achieved sufficient performance on the task\footnote{In the original DAgger the policy rolled out
was stochastically mixed with the supervisor, thus with probability $\beta$ it would either take the supervisor's action
or the robots. The use of this stochastically mix policy was for theoretical analysis. In practice, it is recommended
to set $\beta = 0$ to avoid biasing the sampling~\cite{NIPS2014_5421,ross2010reduction}}.

\subsection{Hierarchy of Supervisors}
Instead of one expert supervisor,we propose a hierarchy  of $M$ supervisors where for each supervisor $\tilde{\pi}_m$, there is an associated expected cumulative reward $R_m = \int \sum^T_{t=1} r(\mathbf{x}_t,\mathbf{u}_t) p(\mathbf{x}|\tilde{\pi}_m)d\mathbf{x})$, where $p(\mathbf{x}|\tilde{\pi}_m)$ denotes the average distribution of states the supervisor encounters. We also denote some measure of cost $C_m$ that is ascribed to the supervisor, such as computational time or monetary expense. We assume that the rank of supervisors in terms of cost is equivalent to the rank of supervisor in the terms of expected cumulative reward. Thus, the cheapest supervisor also receives the lowest expected cumulative reward.

Additionally for two supervisors $\tilde{\pi}_m$ and $\tilde{\pi}_{m+1}$ to be in the hierarchy, they must only disagree in expectation on a subset of the work space to some precision $\tilde{\epsilon}$. Denote the set of states two supervisors disagree as $\mathcal{X}_{m,m+1} = \lbrace \mathbf{x} | ||\tilde{\pi}_m(\mathbf{x}) - \tilde{\pi}_{m+1}(\mathbf{x}) ||^2_2 > \tilde{\epsilon} \rbrace$ .We make this condition because if two supervisors never provided the same examples then all states would have to be relabeled by the next supervisor in the hierarchy.

We formally define two supervisor's in a hierarchy as follows

\begin{definition} Two supervisors $\tilde{\pi}_m$ and $\tilde{\pi}_{m+1}$ are in a hierarchy if  $C_m < C_{m+1}$ , $R_m < R_{m+1}$ and  $\mathcal{X}_{m,m+1} \subset \mathcal{X}$
\end{definition}

We note that a two hierarchical supervisors according to this definition also implies that supervisor $\tilde{\pi}_{m+1}$ in the hierarchy is higher than the set of supervisors $\lbrace \tilde{\pi}_i | \forall i \in [0,m] \rbrace$.

\subsection{A Gambler's DAgger}
Our algorithm is as follows: first iterate through Step 1 and 2 of DAgger for a given supervisor, however after iteration $K$ or when the current policy, $\pi_{\theta_{m,k}}$ is able to achieve a loss $\epsilon$ on the sampled $J$ states at that current iteration (i.e):

$$\frac{1}{J} \sum^J_{j=1} ||\pi_{\theta_{m,k}}(\mathbf{x}_j) - \tilde{\pi}_m(\mathbf{x}_j) ||^2_2 \leq \epsilon$$

Then iterate to the next supervisor in the hierarchy or $m = m+1$. An issue arises though when performing this iterations because now the current dataset $\mathcal{D}_K$ has examples from a different supervisor that receives a smaller expected cumulative reward. This could cause a learning algorithm to try and regress to both labels and diverge from the intended performance \cite{scholkopf2002learning}.

We are thus interested in not storing the dataset collected from the previous supervisor and only remembering the current $\theta_{m,K}$ weight parameters. Thus, we propose only performing DAgger for each individual supervisor in the hierarchy and using the resulting weight vector to bootstrap the learning of the next supervisor. \mlnote{the last step about the dataset is something I want to empirically validate}

@@

\includegraphics[width=\textwidth]{f_figs/Labeling_GUI.pdf}

\caption{\footnotesize  The interface AMT workers see for providing feedback to the robot for the grasping in clutter task. The pink overlay indicates the desire change in the robot position with respect to the current robot state. Then AMT workers can use their intuition for how objects respond to force to provide examples of how the robot should behave. Each window shows a correction in a degree of freedom a) is the rotation of the robots base b) is extending the robot's arm c) is closing the gripper and d) is rotating the turn table}
\vspace*{-20pt}
\label{fig:overlays}
\end{figure*}
